A meta-analysis of mental rotation ability in the first years of life

### 2022-04-19 code/analysis.Rmd
Apply Hedges' small sample correction

```{r, compute_effects, echo=TRUE}
# Add columns to the table of included experiments
included %>%
  mutate(
    # Add difference between condition means
    mean_diff = case_when(
      !is.na(mean_diff) ~ mean_diff,
      TRUE ~ mean_novel - mean_familiar
    ),
    # Add d_z from paired t test of condition means (Rosenthal, 1991)
    d_z_t = t / sqrt(sample_size),
    # Add d_z from ANOVA F value via conversion to a t value
    d_z_f = sqrt(f) / sqrt(sample_size) * sign_f,
    # Add d_z from mean and standard deviation of the difference
    d_z_diff = mean_diff / sd_diff,
    # Add d_av from mean difference and standard devations (assumes r = 0)
    # (Cumming, 2012)
    sd_av = (sd_novel + sd_familiar) / 2,
    d_av = mean_diff / sd_av,
    # Add d from one-sample t test of novelty preference scores
    d_nov_pref = (nov_pref - 0.5) / sd_nov_pref,
    # Choose one type of outcome variable for each experiment
    yi = case_when(
      # 1. If d was reported directed
      !is.na(d) ~ d,
      # 2. If a paired sample t test was reported
      !is.na(d_z_t) ~ d_z_t,
      # 3. If ANOVA was reported
      !is.na(d_z_f) ~ d_z_f,
      # 4. If the difference between means and its SD were reported
      !is.na(d_z_diff) ~ d_z_diff,
      # 5. If the individual condition means and their SDs were reported
      !is.na(d_av) ~ d_av,
      # 6. If a novelty preference score and its SD were reported
      !is.na(d_nov_pref) ~ d_nov_pref
    ),
    # Keep track which type of outcome measure was chosen for each article
    yi_type = case_when(
      !is.na(d) ~ "d",
      !is.na(d_z_t) ~ "d_z_t",
      !is.na(d_z_f) ~ "d_z_f",
      !is.na(d_z_diff) ~ "d_z_diff",
      !is.na(d_av) ~ "d_av",
      !is.na(d_nov_pref) ~ "d_nov_pref",
      TRUE ~ "none"
    ) %>%
      factor(levels = c(
        "d", "d_z_t", "d_z_f", "d_z_diff", "d_av", "d_nov_pref", "none"
      )),
    # Find studies with any value of d_z (from t test, ANOVA, or mean_diff)
    d_z = case_when(
      !is.na(d_z_t) ~ d_z_t,
      !is.na(d_z_f) ~ d_z_f,
      !is.na(d_z_diff) ~ d_z_diff
    ),
    # Compute SD of the difference based on d_z
    sd_z = mean_diff / d_z,
    # Compute correlation based on sd_z and condition SDs
    ri = (sd_z^2 - sd_novel^2 - sd_familiar^2) / (-2 * sd_novel * sd_familiar)
  ) -> dat

# Overview of the different effect sizes
dat %>%
  select(
    article,
    group,
    gender_split,
    yi_type,
    yi,
    d,
    d_z_t,
    d_z_diff,
    d_av,
    d_nov_pref,
    d_z_f,
    ri
  ) %>%
  print(n = Inf)

@@

# Compute standard error of Cohen's d based on assumed correlation
# This will need a sensitivity analysis
r_assumed <- 0.5
dat %>%
  mutate(
    ni = sample_size,
    vi = ((2 * (1 - r_assumed)) / ni) + (yi^2 / (2 * ni)),
    sei = sqrt(vi),
    age_months = age_mean / 30.417,
    age_months_c = age_months - mean(age_months, na.rm = TRUE)
  ) %>%
  filter(!is.na(yi)) %>%
  select(
    article, group, gender_split, age_months_c, female_percent, yi, ri, vi, sei
  ) -> dat_r

@@

```{r, compute_meta, echo=TRUE}
# # Two-level model
# res <- rma(
#   yi, vi,
#   data = dat_r,
#   slab = experiment_ids
# )
# print(res)
# forest(res)

# Three-level model
res_mv <- rma.mv(
  yi, vi,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_mv)
# forest(res_mv)

# # Saving the forest plot
# pdf("forest.pdf", width = 12, height = 12)
# forest(res_mv)
# dev.off()

# # Profile likelihodd plots for checking that REML has converged
# profile(res_mv)

# Meta-regression with age
res_age <- rma.mv(
  yi, vi,
  mods = ~age_months_c,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_age)
# forest(res_age)

# Meta-regression with gender
res_gender <- rma.mv(
  yi, vi,
  mods = ~female_percent,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_gender)
# forest(res_gender)

# Meta-regression with age, gender, and their interaction
res_full <- rma.mv(
  yi, vi,
  mods = ~ age_months_c * female_percent,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_full)
# forest(res_full)
```

@@

# Run prior predictive simulation
res_prior <- brm(
  yi | se(sei) ~ 1 + (1 | article / group),
  data = dat_both_mixed,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 2000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)
summary(res_prior)
# plot(res_prior)

@@

```{r}
table(included$yi_type)
```

>>>>

```{r, compute_effects, echo=TRUE}
# Add columns to the table of included experiments
included %>%
  mutate(
    # Convert mean sample age from days to months and center
    age_months = age_mean / 30.417,
    age_months_c = age_months - mean(age_months, na.rm = TRUE),
    # Add difference between condition means
    mean_diff = case_when(
      !is.na(mean_diff) ~ mean_diff,
      TRUE ~ mean_novel - mean_familiar
    ),
    # Add d_z from paired t test of condition means (Rosenthal, 1991)
    d_z_t = t / sqrt(sample_size),
    # Add d_z from ANOVA F value via conversion to a t value
    d_z_f = sqrt(f) / sqrt(sample_size) * sign_f,
    # Add d_z from mean and standard deviation of the difference
    d_z_diff = mean_diff / sd_diff,
    # Add d_av from mean difference and standard devations (assumes r = 0.5)
    # (Cumming, 2012)
    sd_av = (sd_novel + sd_familiar) / 2,
    d_av = mean_diff / sd_av,
    # Add d from one-sample t test of novelty preference scores
    d_nov_pref = (nov_pref - 0.5) / sd_nov_pref,
    # Choose one type of outcome variable for each experiment
    di = case_when(
      # 1. If d was reported directed
      !is.na(d) ~ d,
      # 2. If a paired sample t test was reported
      !is.na(d_z_t) ~ d_z_t,
      # 3. If ANOVA was reported
      !is.na(d_z_f) ~ d_z_f,
      # 4. If the difference between means and its SD were reported
      !is.na(d_z_diff) ~ d_z_diff,
      # 5. If the individual condition means and their SDs were reported
      !is.na(d_av) ~ d_av,
      # 6. If a novelty preference score and its SD were reported
      !is.na(d_nov_pref) ~ d_nov_pref
    ),
    # Keep track which type of outcome measure was chosen for each article
    di_type = case_when(
      !is.na(d) ~ "d",
      !is.na(d_z_t) ~ "d_z_t",
      !is.na(d_z_f) ~ "d_z_f",
      !is.na(d_z_diff) ~ "d_z_diff",
      !is.na(d_av) ~ "d_av",
      !is.na(d_nov_pref) ~ "d_nov_pref",
      TRUE ~ "none"
    ) %>%
      factor(levels = c(
        "d", "d_z_t", "d_z_f", "d_z_diff", "d_av", "d_nov_pref", "none"
      )),
    # Find studies with any value of d_z (from t test, ANOVA, or mean_diff)
    d_z = case_when(
      !is.na(d_z_t) ~ d_z_t,
      !is.na(d_z_f) ~ d_z_f,
      !is.na(d_z_diff) ~ d_z_diff
    ),
    # Apply small sample correction using Hedges' exact method
    # See http://dx.doi.org/10.20982/tqmp.14.4.p242
    df = 2 * (sample_size - 1),
    j = exp(lgamma(df / 2) - log(sqrt(df / 2)) - lgamma((df - 1) / 2)),
    gi = di * j,
    # Compute empirical correlation based on sd_z and condition SDs
    sd_z = mean_diff / d_z,
    ri = (sd_z^2 - sd_novel^2 - sd_familiar^2) / (-2 * sd_novel * sd_familiar)
  ) -> dat

# Overview of the different effect sizes
dat %>%
  select(
    article,
    group,
    gender_split,
    gi,
    di,
    di_type,
    d,
    d_z_t,
    d_z_f,
    d_z_diff,
    d_av,
    d_nov_pref,
    ri
  ) %>%
  print(n = Inf)

@@

# Compute standard error of Cohen's d based on assumed correlation
# This will need a sensitivity analysis
r_assumed <- 0.5
dat %>%
  mutate(
    ni = sample_size,
    vi = ((2 * (1 - r_assumed)) / ni) + (gi^2 / (2 * ni)),
    sei = sqrt(vi)
  ) %>%
  filter(!is.na(gi)) %>%
  select(
    article, group, gender_split, gi, ni, vi, sei, age_months_c, female_percent
  ) -> dat_r

@@

```{r, compute_meta, echo=TRUE}
# # Two-level model
# res_rma <- rma(
#   gi, vi,
#   data = dat_r,
#   slab = experiment_ids
# )
# print(res_rma)
# forest(res_rma)

# Three-level model
res_ml <- rma.mv(
  gi, vi,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_ml)
# forest(res_ml)

# # Saving the forest plot
# pdf("forest.pdf", width = 12, height = 12)
# forest(res_ml)
# dev.off()

# # Profile likelihodd plots for checking that REML has converged
# profile(res_ml)

# Meta-regression with age
res_age <- rma.mv(
  gi, vi,
  mods = ~age_months_c,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_age)
# forest(res_age)

# Meta-regression with gender
res_gender <- rma.mv(
  gi, vi,
  mods = ~female_percent,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_gender)
# forest(res_gender)

# Meta-regression with age, gender, and their interaction
res_full <- rma.mv(
  gi, vi,
  mods = ~ age_months_c * female_percent,
  random = ~ 1 | article / group,
  data = dat_both_mixed,
  slab = paste(article, group, sep = ", ")
)
print(res_full)
# forest(res_full)
```

@@

# Run prior predictive simulation
res_prior <- brm(
  gi | se(sei) ~ 1 + (1 | article / group),
  data = dat_both_mixed,
  prior = prior_c,
  sample_prior = "only",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 2000,
  cores = 4,
  control = list(adapt_delta = 0.99)
)
summary(res_prior)
# plot(res_prior)

@@

```{r}
table(included$di_type)
```
